---
title: "Final PCR and PLS"
author: "Owen R. Evans"
date: "8/25/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using dietary data from 2013-2014 NHANES.  Need demographic and depression data for the same cohort. 

``` {r Data Processing}
library(rio)
library(tidyverse)
setwd("~/Desktop/Predictive Modeling Final Project")
depr.df <-import("DPQ_H_2014.XPT") # PHQ-09 Data, Severity of Depression

library(sticky) # maintain attribute labels
depr.df <- select(depr.df, -11)
depr.df <- na.omit(depr.df)
depr.df$DepressionScore = rowSums(depr.df[,-1])
depr.df <- filter(depr.df, DepressionScore<27)
depr.df<- select(depr.df, SEQN, DepressionScore) # Total PHQ-09 Score

# Import quantitative dietary data
# Food nutrients per day - multiple entries per day
# Need to aggregate nutrient intake per SEQN

diet.df <-import("DR1IFF_H_2014.XPT")
diet.df <- select(diet.df, SEQN, 20:84)
sticky_all(diet.df)

# aggregate daily dietary intake for each unique SEQN/observation
unique.seqn <- diet.df$SEQN
sticky_all(diet.df)
diet_agg.df <- aggregate(diet.df, data.frame(unique.seqn), sum)
diet_agg.df$SEQN <- NULL

# Import demographic data
demo.df <-import("DEMO_H_2014.XPT") # Demographics
demo.df <- sticky_all(demo.df)
demo.df <- select(demo.df, SEQN, RIAGENDR, RIDAGEYR)

# Merge demographic with depression scores
sticky_all(demo.df)
master_depr.df <- merge(depr.df, demo.df, by="SEQN")
summary(master_depr.df)

# filter master for males, aged 18-40
# merge with dietary data
# scale predictors 
master_depr.males.df <- filter(master_depr.df, RIAGENDR==1)
master_depr.males.df <- filter(master_depr.males.df, 
                               RIDAGEYR >= 20 & RIDAGEYR <=40)
master_diet_males.df <- merge(master_depr.males.df, diet_agg.df, by="SEQN")
master_diet_males.df <- na.omit(master_diet_males.df)
master_diet_males.df <- subset(master_diet_males.df, 
                               select=-c(RIAGENDR, RIDAGEYR, SEQN))
master_male_scaled.df <- as.data.frame(scale(master_diet_males.df[,-2]))

# filter master for females, aged 18-40
# merge with dietary data
# scale predictors
master_depr.females.df <- filter(master_depr.df, RIAGENDR==2)
master_depr.females.df <- filter(master_depr.females.df, 
                               RIDAGEYR >= 20 & RIDAGEYR <=40)
master_diet_females.df <- 
  merge(master_depr.females.df, diet_agg.df, by="SEQN")
master_diet_females.df <- na.omit(master_diet_females.df)
master_diet_females.df <- subset(master_diet_females.df, 
                               select=-c(RIAGENDR, RIDAGEYR, SEQN))
master_female_scaled.df <- as.data.frame(scale(master_diet_females.df[,-2]))



```

``` {r PCR Fit, Males}
# NULL Model MSE
mean_male.y <- mean(master_male_scaled.df$DepressionScore)
null.MSE.male <- mean((master_male_scaled.df $DepressionScore-mean_male.y)^2)

# OLS - standard linear fit
lm.fit.male <- lm(DepressionScore~., data=master_male_scaled.df)
summary(lm.fit.male)

#pcr fit for male subgroup 
library(pls)

set.seed(1234)
train.m <- sample(1:nrow(master_male_scaled.df ),
                  0.5*nrow(master_male_scaled.df ))
test.m <- -train.m

# fit pcr model for training data
pcr.fit.males<- pcr(DepressionScore~., 
               data=master_male_scaled.df [train.m,], 
               scale=TRUE, validation="CV")

summary (pcr.fit.males)
validationplot(pcr.fit.males, val.type="MSEP")

# Calculate Test Error for 15 components
pcr.pred.male <- predict(pcr.fit.males, 
                         newdata = master_male_scaled.df [test.m,], ncomp=15)

pcr.test.MSE.males <- mean((pcr.pred.male-master_male_scaled.df $DepressionScore[test.m])^2)
pcr.test.MSE.males
# high mean squared error compared to OLS, ridge or lasso.
# same as null model - diet has no explanatory power

# Fit Full Model
# Plot Actual vs. Predicted
pcr.fit.full.males <- pcr(DepressionScore ~., 
                    data=master_male_scaled.df , ncomp = 19, scale=TRUE)
pcr.preds.full.males <- predict(pcr.fit.full.males, newdata = master_male_scaled.df , ncomp=19)
pcr.full.MSE.males <- 
  mean((pcr.preds.full.males-master_diet_males.df $DepressionScore)^2)

#17.7 MSE
plot(pcr.fit.full.males)

```

``` {r PCR Fit, females}
# NULL Model MSE
mean_female.y <- mean(master_diet_females.df$DepressionScore)
Null.MSE.Females <- mean((master_diet_females.df$DepressionScore-mean_female.y)^2)

#pcr fit for female subgroup 
library(pls)

set.seed(1234)
train.f <- sample(1:nrow(master_diet_females.df),
                  0.5*nrow(master_diet_females.df))
test.f <- -train.f

# fit pcr model for training data
pcr.fit.females<- pcr(DepressionScore~., 
               data=master_diet_females.df[train.f,], 
               scale=TRUE, validation="CV")

summary (pcr.fit.females)
validationplot(pcr.fit.females, val.type="MSEP")

# Calculate Test Error for 3 components
pcr.pred.female <- predict(pcr.fit.females, 
                         newdata = master_diet_females.df[test.f,], ncomp=3)

pcr.test.MSE.females <- mean((pcr.pred.female-master_diet_females.df$DepressionScore[test.f])^2)
pcr.test.MSE.females
# high mean squared error compared to OLS, ridge or lasso.
# same as null model - diet has no explanatory power

# Fit Full Model
# Plot Actual vs. Predicted
pcr.fit.full.females <- pcr(DepressionScore ~., 
                    data=master_diet_females.df, ncomp = 4, scale=TRUE)
pcr.preds.full.females <- predict(pcr.fit.full.females, newdata = master_diet_females.df, ncomp=4)
pcr.full.MSE.females <- 
  mean((pcr.preds.full.females-master_diet_females.df$DepressionScore)^2)

#17.7 MSE
plot(pcr.fit.full.females)

```

Efforts to model the severity of depression symptoms as a function of dietary intake via principal components regression or partial least squares were largely unsuccessful.  The models obtained did not improve predictive accuracy relative to that observed for a standard linear fit and that observed solely for the NULL model.   This highlights the difficulty in developing predictive models for human behavior and mental health.   Additionally, despite best attempts, the presence of several confounding variables (geography, race, gender, age, health issues, income) can significantly affect the predictive accuracy of such a model.  

To demonstrate the utility of both PCR and PLS,  a new problem was explored.  Specifically,  the NHANES extensive data set on dietary intake can allow one to predict and determine the relationship between dietary intake and body mass index.  Both PCR and PLS are ideal for this effort, as the dataset contains a wealth of continuous and potentially collinear variables that may ultimately complicate models based upon standard least squares regression. 

``` {r BMI vs. Dietary}

# need BMI data
body.df <- import("BMX_H.XPT")
body.df <- sticky_all(body.df)
body.df <- dplyr::select(body.df, SEQN, BMXBMI) 
body.df <- na.omit(body.df)

# Import quantitative dietary data
# Food nutrients per day - multiple entries per day
# Need to aggregate nutrient intake per SEQN

diet.df <-import("DR1IFF_H_2014.XPT")
diet.df <- dplyr::select(diet.df, SEQN, 20:84)
sticky_all(diet.df)

# aggregate daily dietary intake for each unique SEQN/observation
SEQN <- diet.df$SEQN
diet.df$SEQ1 <-diet.df$SEQN
sticky_all(diet.df)
diet_agg.df <- aggregate(diet.df, data.frame(SEQN), sum)
diet_agg.df[,2] <- NULL
diet_agg.df <- na.omit(diet_agg.df)

# Import demographic data
demo.df <-import("DEMO_H_2014.XPT") # Demographics
demo.df <- sticky_all(demo.df)
demo.df <- dplyr::select(demo.df, SEQN, RIAGENDR, RIDAGEYR)

# Merge with Dietary Data
master_diet.df <- merge(body.df, diet_agg.df, by="SEQN")
master_diet.df <- dplyr::select(master_diet.df, -c(1))
master_diet.df <- na.omit(master_diet.df)

```

``` {r PCR}
#df.scaled <- as.data.frame(scale(master_diet.df))
library(pls)
set.seed(1234)
train.f <- sample(1:nrow(master_diet.df),0.5*nrow(master_diet.df))
test.f <- -train.f

# fit pcr model to training data
pcr.fit.bmi<- pcr(BMXBMI~., 
               data=master_diet.df[train.f,], 
               scale=TRUE, validation="CV")

summary (pcr.fit.bmi)
validationplot(pcr.fit.bmi, val.type="MSEP")

# Calculate Test Error for 28 components
pcr.pred.bmi <- predict(pcr.fit.bmi, 
                         newdata = master_diet.df[test.f,], ncomp=28)

pcr.test.MSE <- mean((pcr.pred.bmi-master_diet.df$BMXBMI[test.f])^2)
pcr.test.MSE # A little better than NULL model

pcr.null <- mean((mean(master_diet.df$BMXBMI)-master_diet.df$BMXBMI)^2)
pcr.null # A little better

```
``` {r PLS}
library(pls)
set.seed(1234)
train.g <- sample(1:nrow(master_diet.df),
                  0.5*nrow(master_diet.df))
test.g <- -train.g

# fit pcr model for training data
pls.fit<- plsr(BMXBMI~., 
               data=master_diet.df[train.g,], 
               scale=TRUE, validation="CV")

summary (pls.fit)
validationplot(pls.fit, val.type="MSEP")

# Calculate Test Error for 5 components
pls.pred <- predict(pls.fit, 
                         newdata = master_diet.df[test.g,], ncomp=5)

pls.test.MSE <- mean((pls.pred-master_diet.df$BMXBMI[test.g])^2)
pls.test.MSE

# Fit Full Model
# Plot Actual vs. Predicted
pls.fit.full<- plsr(BMXBMI ~., 
                    data=master_diet.df, ncomp = 5, scale=TRUE)
pls.preds.full <- predict(pls.fit.full, 
                          newdata = master_diet.df, ncomp=5, scale=TRUE)
pls.full.MSE<- 
  mean((pls.preds.full-master_diet.df$BMXBMI)^2)
pls.full.MSE

plot(pls.fit.full)
abline(0,1, col="red")

plot(pcr.fit.bmi)
abline(0,1, col="red")

```




